---
title: "RWorksheet_Sorenio#4b"
output: pdf_document
date: "2024-10-28"
---

```{r}
# 1
vectA <- c(1, 2, 3, 4, 5)

matrixB <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixB[i, j] <- abs(vectA[i] - vectA[j])
  }
}
matrixB
```
```{r}
# 2
for (i in 1:5) {
  for (j in 1:i) {
    cat("* ")
  }
  cat("\n")
}
```

```{r}
# 3
start1 <- as.integer(readline(prompt="Enter the first number of the Fibonacci sequence: "))
start2 <- as.integer(readline(prompt="Enter the second number of the Fibonacci sequence: "))

fibonacci <- c(start1, start2)
repeat {
  next_value <- tail(fibonacci, 2)[1] + tail(fibonacci, 2)[2]
  if (next_value > 500) break
  fibonacci <- c(fibonacci, next_value)
}
print(fibonacci)
```
```{r}
# 4
data <- read_csv("C:/Users/User/Documents/sample_data.csv")

data
```
```{r}
# 4 a
data <- read_csv("C:/Users/User/Documents/sample_data.csv", show_col_types = FALSE)
print(head(data, 6))
```
```{r}
# 4b
fem <- subset(data, Gender == "F")
male <- subset(data, Gender == "M")
cat("Female count:", nrow(fem),"\n")
cat("Male count:", nrow(male),"\n")
```
```{r}
# 4c
gencount <- table(data$Gender)
barplot(gencount, main= "Gender Distribution", col = c("red", "blue"), 
        xlab="Gender", ylab="Count", legend=TRUE)
```
```{r}
# 5 a
library(ggplot2)

categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
spending <- c(60, 10, 5, 25)


data <- data.frame(categories, spending)


data$percentage <- data$spending / sum(data$spending) * 100


ggplot(data, aes(x = "", y = percentage, fill = categories)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(round(percentage), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(title = "Monthly Income Distribution of Dela Cruz Family") +
  theme_void() +
  scale_fill_brewer(palette = "Set1")
```

```{r}
# 6 a
data(iris)
str(iris)

```
```{r}
# 6 b
means <- colMeans(iris[, 1:4])
means
```
```{r}
# 6 c
specscounts <- table(iris$Species)
pie(specscounts, 
    main = "Species Distribution in Iris Dataset", 
    col = c("salmon", "purple", "lightpink"), 
    labels = paste(names(specscounts), specscounts))
legend("topright", legend = names(specscounts), fill = c("salmon", "purple", "lightpink"))
```
```{r}
# 6 d
setosa <- iris[iris$Species == "setosa", ]
versicolor <- iris[iris$Species == "versicolor", ]
virginica <- iris[iris$Species == "virginica", ]

tail(setosa)
tail(versicolor)
tail(virginica)
```

```{r}
# 6 e
iris$Species <- as.factor(iris$Species)

plot(iris$Sepal.Length, iris$Sepal.Width, 
     col = iris$Species, 
     pch = 19, 
     main = "Iris Dataset", 
     sub = "Sepal Width and Length", 
     xlab = "Sepal Length", 
     ylab = "Sepal Width")
legend("topright", legend = levels(iris$Species), col = 1:3, pch = 19)
```
# 6 f. Interpretation
# In the analysis of the iris dataset, we found that it has 150 samples and five variables: sepal length, sepal width, petal length, petal width, and species. The mean sepal length is about 5.84 cm, which tells us the average size of the flowers.

# The created pie chart shows the distribution of species, revealing that setosa is the most common. We also looked at the last six rows of each species to see their specific measurements.

# Finally, the scatterplot of sepal length versus sepal width showed that setosa flowers generally have shorter sepals compared to versicolor and virginica, which are more similar. Overall, this analysis helps us understand the differences among the iris species.

```{r}
# 7

library(readxl)

alexa_data <- read_excel("C:\\Users\\User\\Downloads\\alexa_file.xlsx")
alexa_data
```







